Contracts for Contracts

### 2023-06-01 paper/introduction.tex
some intro

Indeed, in recent years, we have seen a number of failures of smart contract
systems due to attacks utilizing vulnerable contract programs.
For example, in the famous DAO hack\note{cite}, the attacker is able to utilize the
callback mechanism to create recursive function calls and temporarily stole
money, causing a total lose of \$50 million.
The root cause of this attack is that the contract does not deploy defensive
checks before performing critical actions.

@@

We observe that popular smart contract languages (such as Solidity) do
not provide \emph{rich, effective, convenient} means to specify and enforce
contract behaviors.
Simple first-order assertion primitives (e.g., @require@) are indeed common,
but they cannot effectively check callable addresses or function values, which
are a common pattern used in smart contract programs.
Moreover, many of the critical condition in smart contracts are \emph{temporal} --
the order of happened events matters.
For instance, the defense to the DAO attack \note{cite} is to
prevent ``reentrancy'' via external functions, which requires that the function
of interest cannot be invoked \emph{before} the current call of this function is finished.

The lack of sufficient linguistic means to specify and enforce subtle but
critical behaviors discourages programmers to use concise higher-level
abstractions when writing smart contract code.
Instead, to be secure, programmers are obliged to write verbose code with
lower-level or no abstractions, which conflates the main business logic with
behavioral checks, further leading to poor readability and maintainability.
For example, a common way to implement ``reentrancy guards'' is to manually
make an additional boolean state, marking if the function is finished.
And often worse, defensive checks can be easily forgotten or disregarded,
leading to vulnerable code causing real monetary lose.

%Apart from the investigating effort to certify and verify smart contract
%programs, language designers are responsible to empower programmers with better
%linguistic solutions to secure smart contracts in the first place.
In this paper, we develop practical \emph{behavioral
software contracts} that provide a lightweight solution to secure
\emph{smart contracts}.
In particular, we develop a specification language and monitor for higher-order
and temporal behaviors in the Solidity language, demonstrating that they
are an effective approach to

\subsection*{\textbf{Behavioral Contracts for Solidty}}

Similar to smart contracts, behavioral contracts is an agreement too.  They
specify the assumptions and guarantees between software components.
In languages with behavioral contracts, the specified assumptions and
guarantees will be monitored and reported when violated.
Pioneered by the Eiffel language \cite{DBLP:books/ph/Meyer91} and the
Design-by-Contract methodology \cite{DBLP:conf/tools/Meyer98a}, behavioral
contracts have been widely adopted in Java, Haskell, C\#, Python, Racket,
Elixir \cite{DBLP:conf/erlang/0001BBHMEF22}, etc. \note{cite}

Solidity is a programming language to build smart contract programs that run on
the Ethereum platform.
As one of the most popular languages for smart contracts, Solidity has been
used \todo{to do cool things}.
However, Solidity and its running platform Ethereum is a complex system.
Carelessly written contracts can cause serious bugs and loses.
We identify \todo{solidity's tricky feature}

To build robust smart contract programs, we design and implement a practical
extension of Solidity, namely \lang.
The extension enriches the Solidity language with \emph{higher-order} and
\emph{temporal} behavioral contracts, which are absent in existing approaches.
With higher-order and temporal behavioral specification, many downstream
tasks become further amenable, including but not limited to static verification,
guided testing, or specification extraction.

\subsection*{\textbf{Contracts for Callable Addresses and Higher-Order Functions}}

In Solidity, callbacks can be in two forms, addresses and higher-order functions.

\todo{restrict possible call targets}
\todo{why static verification fails}
\zz{address vs. function call?}

>>>>

Indeed, in recent years, we have seen a number of failures of smart contract
systems due to attacks utilizing vulnerable contract programs.
For example, in the famous DAO hack\note{cite}, the unauthorized attacker is
able transfer money by constructing recursive function calls with
the callback mechanism, causing a total lose of \$50 million.
The root cause of this attack is that the deployed contract does not
checks reentrancy defensively before performing critical actions.

@@

To prevent such attacks, the programmer needs to specify those critical
conditions. The execution is allowed when those conditions are met.
However, we observe that existing popular smart contract languages (e.g.
Solidity) do not provide \emph{rich, effective, convenient} means to specify
and enforce contract behaviors.
Simple first-order assertion primitives (e.g., @require@) are indeed available,
but they cannot effectively examine callable addresses or function values, which
are a common pattern used in smart contract programs.
Moreover, many of the critical condition in smart contracts are \emph{temporal}
-- the order of happened events matters.
For instance, to defense the DAO attack caused by reentrancy via
external functions, it requires to check that the function of interest cannot
be invoked \emph{before} the current call of the same function is finished.

The lack of sufficient linguistic means to specify and enforce subtle but
critical behaviors discourages programmers to use concise higher-level
abstractions when writing smart contract code.
Instead, to be secure, programmers are obliged to write verbose code with
lower-level or no abstractions, which conflates the main business logic with
behavioral checks, further leading to poor readability and maintainability.
Taking the DAO attack as example again, a common way to implement ``reentrancy
guards'' is to manually define an additional boolean state, indicating
if the function is finished.
And often worse, defensive checks can be easily forgotten or disregarded,
leading to vulnerable code causing real monetary lose.

%Apart from the investigating effort to certify and verify smart contract
%programs, language designers are responsible to empower programmers with better
%linguistic solutions to secure smart contracts in the first place.
In this paper, we argue that \emph{behavioral software contracts}
\cite{DBLP:conf/tools/Meyer98a} should play a fundamental role in the
development of \emph{smart contract} programs.  Similar to smart contracts,
behavioral contracts is an agreement too. They specify the assumptions and
guarantees between software components. The specified assumptions and
guarantees will be monitored and reported when violated in languages with
behavioral contracts.

Pioneered by the Eiffel language \cite{DBLP:books/ph/Meyer91} and the
design-by-contract methodology \cite{DBLP:conf/tools/Meyer98a}, various styles
of behavioral contracts have been widely adopted in Java, Haskell, C\#, Python,
Racket, Elixir \cite{DBLP:conf/erlang/0001BBHMEF22}, etc. \note{cite}
In this paper, we develop a behavioral contract system \lang for the Solidity
language, providing a practical specification and monitoring system
for both higher-order and temporal behaviors.
Solidity is a programming language to build smart contract programs that run on
the Ethereum platform.  As one of the most popular languages for smart
contracts, Solidity has been used \todo{to do cool things}.

We demonstrate that behavioral contracts are an effective approach
to improve the robustness and quality of smart contract programs.
\todo{summarize what we can prevent/achieve}

%\lang enriches the Solidity language with \emph{higher-order} and
%\emph{temporal} behavioral contracts, which are absent in existing approaches.
%With higher-order and temporal behavioral specification, many downstream
%tasks become further amenable, including but not limited to static verification,
%guided testing, or specification extraction.

%However, Solidity and its running platform Ethereum is a complex system.
%Carelessly written contracts can cause serious bugs and loses.
%We identify \todo{solidity's tricky feature}

\subsection*{\textbf{Contracts for Higher-Order Values}}

One type of values in Solidity that exhibits higher-order behaviors is \emph{addresses}.

Functions in Solidity are first-class too, meaning that functions can be used as
arguments for and returned from functions. However, at this moment
\footnote{Solidity version 8.20} Solidity has not supported writing anonymous
functions (lambda expressions) or nested (named) functions.
Only explicitly defined top-level functions can be used in first-class ways,
which discourages programmers to use higher-order functions due to its
inconvenience and verbosity.
Moreover, this restriction poses challenges to seamless implement first-class
behavioral contract and monitor system (e.g. as in \cite{DBLP:conf/icfp/FindlerF02})
without an expensive whole program transformation.
Therefore, in this paper we focus on the contract and monitoring of addresses values
and leave monitoring for higher-order functions as future work when
Solidity has a proper support for lambda expressions.

\todo{restrict possible call targets}
\todo{why static verification fails}
